---
title: "Pharmacogenomics Next Generation Sequencing Run Quality Control Metrics"
author: "Galenvs Sciences Inc."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2:
    keep_tex: true
    highlight: monochrome
    toc: false
    toc_depth: 2
    number_sections: true
    fig_caption: true
params:
  Barcode_summary_path: ""
  Amplicon_summary_path: ""
  depth_files_folder_path: ""    
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
  \usepackage[format=plain,labelfont={bf,it},textfont={}]{caption}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{lscape}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(ggpubr)
library(knitr)
library(kableExtra)
library(scales)
library(viridis)
library(xtable)
library(purrr)
library(gplots)
#qiacuityData <- "~/Documents/galenvs/cfDNA/qiacuity/data/2023-04-13"
# library(formatR)
# opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/jonat/Desktop/galenvs_dashboard/server/records/") 
# knitr::opts_knit$set(root.dir = "/home/testvm/galenvs_dashboard/server/records/") 
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE,  class.output="shadebox")
```


```{r data import}
#params<- list(
#  Barcode_summary_path = "GalenvsPGx_Thermo_v2_Samples_1_to_14_torrent-server_67.bc_summary.xls",
# Amplicon_summary_path = " ",
#  depth_files_folder_path = "depth/"
#  )
# params = list(Barcode_summary_path = 'experiment_data_140723_1108_day second /barcodeSummary_01_140723_1108_day second .xls', Amplicon_summary_path = 'experiment_data_140723_1108_day second /ampliconSummary_01_140723_1108_day second .xls', depth_files_folder_path = 'experiment_data_140723_1108_day second /depth_140723_1108_day second /')

# Import Barcode Summary File
bsf <- read_delim(params$Barcode_summary_path)

# Import Depth Files
files_to_read <- list.files(path = params$depth_files_folder_path, pattern = "\\.xls$", full.names = T)
all_files <- lapply(files_to_read, function(x) {
   read.table(file = x, sep = '\t', header = TRUE)
})
names(all_files) <- files_to_read %>% gsub(".*//", "", .)

# Import Amplicon Summary File
amplicon <- read_delim(params$Amplicon_summary_path)
```

```{r newColumns}
files2 <-
          all_files %>% 
            lapply(function(x) 
            cbind(x, len = x$contig_end - x$contig_srt + 1)) %>%
            lapply(function(x) 
            cbind(x, rat500 = x$cov500x/x$len)) %>%  
            lapply(function(x) 
            cbind(x, rat2500 = x$cov2500x/x$len)) 

```

```{r bindRows}
filesBind <- bind_rows(files2, .id="file")
```

```{r createSummary}
summaryTibble <- bsf %>% dplyr::select(`Barcode ID`, `Sample Name`, `Mean Depth`, `Uniformity`)
```


```{r BaseDepthAnalysis}
# Pass / Fail on Mean Depth
    summaryTibble$`Depth QC Check` <- "FAIL"
    summaryTibble$`Depth QC Check`[summaryTibble$ `Mean Depth` > 2500] <- "PASS"
    
# Which regions fail (amplicons that have less than 500X by sample)
    baseDepthLess500 <- lapply(files2, function(x) filter(x, rat500 <1)) %>% lapply(function(x) select(x, region_id))
```


```{r UniformityAnalysis}
# Pass / Fail on Mean Depth
    summaryTibble$`Uniformity QC Check` <- "FAIL"
    summaryTibble$`Uniformity QC Check`[summaryTibble$ `Uniformity` > 93] <- "ACCEPTABLE"
    summaryTibble$`Uniformity QC Check`[summaryTibble$ `Uniformity` > 96] <- "OPTIMAL"
```

```{r AmpliconCoverageAnalysis}
    longAmplicon <- amplicon %>% pivot_longer(cols = 3:ncol(amplicon))
    coverageFreq <- longAmplicon %>% 
      filter(value<500) %>% 
      dplyr::select(name) %>% 
      table() %>%
      `/` (nrow(amplicon)) %>% 
      as.data.frame()
    
    colnames(coverageFreq)[1] <- "Barcode ID"
    coverageFreq$`Coverage QC Check` <- "FAIL"
    coverageFreq$`Coverage QC Check`[coverageFreq$Freq < 0.05] <- "PASS"
    
    summaryTibble <- summaryTibble %>% full_join(coverageFreq[,c(1,3)])

```

```{r overallQC}
summaryTibble <- summaryTibble %>% mutate(`Overall QC Checkpoint` = if_else(`Depth QC Check` == "PASS" & `Uniformity QC Check` == "ACCEPTABLE" | `Uniformity QC Check` == "ACCEPTABLE" & `Coverage QC Check` == "PASS", "PASS", "FAIL"))
```


# Summary

```{r}
summaryColours <- summaryTibble$`Overall QC Checkpoint` %>% gsub("PASS", "olive", .) %>% gsub("FAIL", "red", .)

summaryTibble %>% 
  dplyr::select(`Barcode ID`, `Sample Name`, `Depth QC Check`, `Uniformity QC Check`, `Coverage QC Check`, `Overall QC Checkpoint`) %>% 
  sapply(function(x){str_wrap(x, 10)}) %>% 
  kbl(booktabs=TRUE, longtable=TRUE, caption = "Summary of QC check results.")  %>% 
  column_spec(1:6, width = paste0(c(2,2,1.6,3,1.6,1), "cm")) %>% 
  column_spec(6, color=summaryColours)
```
\newpage

\tableofcontents

\newpage
# Base Depth
## Mean Depth
The mean depth for each sample is shown in ***table \@ref(tab:meanDepthResults)***.

```{r meanDepthResults}
summaryTibble %>% 
  dplyr::select(`Barcode ID`, `Sample Name`, `Mean Depth`, `Depth QC Check`) %>% 
  kbl(booktabs=TRUE, longtable=TRUE, caption = "Mean Depth Results.")
```

## Regions with <100% Bases Read 500X (by Sample)

The distribution of the fraction of bases read 500X in a given amplicon is shown in ***figure \@ref(fig:depthHeatmap)***.

```{r depthHeatmap, fig.cap="Heatmap showing the distribution of the fraction of bases read 500X in a given amplicon. Rows represent amplicons and columns represent samples", fig.height=12, fig.width=12, fig.scap= "Depth 500X Heatmap"}
wideFiles <- filesBind %>% select(file, region_id, rat500) %>% pivot_wider(names_from = file, values_from = rat500) %>% as.data.frame()
rownames(wideFiles) <- wideFiles$region_id
wideMatrix <- wideFiles[,2:ncol(wideFiles)] %>% as.matrix

wideMatrix %>% heatmap.2(col=turbo,trace="none", labRow = "")
```


The following regions do not meet the criteria of having 100% of bases read 500X (by sample). The total number of unique regions that do not meet this criteria across samples is `r baseDepthLess500 %>% lapply(function(x){as.vector(x)}) %>% unlist() %>% unique() %>% length()` regions.

```{r}
baseDepthLess500 %>% lapply(function(x){as.vector(x)})
```

## Intersect Between Regions with <100% Bases Read 500X in at least Half of the Samples

The following `r baseDepthLess500 %>% lapply(function(x){as.vector(x)}) %>% unlist() %>% unique() %>% length()` regions have <100% bases read 500X in at least one sample.

```{r}
baseDepthLess500 %>% lapply(function(x){as.vector(x)}) %>% unlist() %>% unique()
```


```{r}

halfSamples <- length(baseDepthLess500) /2

depthFlagOverall <- baseDepthLess500 %>% lapply(function(x){as.vector(x)}) %>% unlist() %>% table()
depthFlagIntersectHalf <- depthFlagOverall[depthFlagOverall > halfSamples] %>% names()

```

The following `r depthFlagIntersectHalf %>% length()` regions have <100% bases read 500X in at least half of the samples.

```{r}
depthFlagIntersectHalf
```

\newpage

# Amplicon Coverage

## Coverage Frequency

The fraction of amplicons in each sample that are not covered at least 500X is shown in ***table \@ref(tab:coverageFrequencyResults)***.

```{r coverageFrequencyResults}
summaryTibble %>% full_join(coverageFreq) %>% 
  dplyr::select(`Barcode ID`, `Sample Name`, `Freq`, `Coverage QC Check`) %>% 
  kbl(booktabs=TRUE, longtable=TRUE,   caption = "Coverage Frequency Results. `Freq` is the fraction of amplicons that are not covered at least 500X.")
```

The distribution of the coverage of amplicons is shown in ***figure \@ref(fig:coverageHeatmap)***.

```{r coverageHeatmap, fig.cap="Heatmap showing the distribution of the coverage of each amplicon. Rows represent amplicons and columns represent samples", fig.height=12, fig.width=12, fig.scap= "Coverage Heatmap"}

ampMatrix <- amplicon[3:ncol(amplicon)] %>% as.matrix()
rownames(ampMatrix) <- amplicon$Target

ampMatrix  %>% heatmap.2(col=turbo,trace="none", labRow = "")
```


## Intersect Between Regions covered <500X in at least Half of the Samples

The following amplicons were covered less than 500X in at least half of the samples.

```{r}
amplicon2 <- amplicon
amplicon2[,3:ncol(amplicon2)] <- amplicon[,3:ncol(amplicon)] > 500
amplicon2[,3:ncol(amplicon2)] <- amplicon2[,3:ncol(amplicon2)] %>% sapply(function(x){as.integer(x)})
amplicon2$rowSums <- rowSums(amplicon2[,3:ncol(amplicon2)])

amplicon2 %>% dplyr::filter(rowSums<halfSamples) %>% dplyr::select(Target) %>% as.vector()
```

\newpage

# Uniformity

The uniformity results for each sample are shown in ***table \@ref(tab:uniformityTable)***.


```{r uniformityTable}
summaryTibble %>% 
  dplyr::select(`Barcode ID`, `Sample Name`, `Uniformity`, `Uniformity QC Check`)  %>% 
  kbl(booktabs=TRUE, longtable=TRUE, caption = "Uniformity Results.")
```


